---
title: "curate_reference_feature_dataset"
author: "Kyle Gontjes"
date: "`r Sys.Date()`"
output: html_document
---
 
```{r setup, include=FALSE}
require(knitr)
opts_knit$set(echo = TRUE,warning = FALSE, message = FALSE) 
rm(list = ls())
opts_knit$set(root.dir='~/Desktop/gl_mount/Analysis/Combination_resistance/pre-intro-blbli-resistance-ms/')
```
 
```{r}
library(genbankr)
library(tidyverse)

# Print
Sys.info()
sessionInfo() 
```
 
```{r}
kpnih1 <- readGenBank(GBAccession("NZ_CP008827"),partial=TRUE)
# Pull Features Using Locus Tag
kpnih1_features <- cds(kpnih1) %>% as.data.frame
  # Formatting Genome set
  kpnih1_features_final <- kpnih1_features %>% dplyr::select(locus_tag,start,end,strand,gene,product,protein_id) %>% 
    mutate(length = end - start,
           locus_tag = make.unique(locus_tag)) 
saveRDS(kpnih1_features_final,"./data/references/KPNIH1/KPNIH1_features.RDS")

KPC_plasmid <- readGenBank(GBAccession("NZ_CP008830"),partial=TRUE) 
# Pull Features Using Locus Tag
KPC_plasmid_features <- cds(KPC_plasmid) %>% as.data.frame
  # Formatting Genome set
  KPC_plasmid_features_final <- KPC_plasmid_features %>% dplyr::select(locus_tag,start,end,strand,gene,product,protein_id) %>% 
    mutate(length = end - start,
           locus_tag = make.unique(locus_tag)) 
   
saveRDS(KPC_plasmid_features_final,"./data/references/KPNIH1_plasmid_pKpQIL6e6/KPNIH1_KPC_Plasmid_features.RDS")
