---
title: "S Figure 9: KPC Copy Number"
author: "Kyle Gontjes"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
require(knitr)
opts_knit$set(echo = T,warning = FALSE, message = FALSE,error=F,comment=NA)
opts_chunk$set(echo = T,warning = FALSE, message = FALSE,error=F,comment=NA)
opts_knit$set(root.dir='~/Desktop/gl_mount/Analysis/Combination_resistance/pre-intro-blbli-resistance-ms/')
```

# Environment
```{r}
packages <- c("tidyverse","cowplot","ggtree","ape","phytools","gridExtra","kableExtra","colorRamp2","RColorBrewer","ComplexHeatmap",'phyloAMR','ggtext')

# Load packages
invisible(lapply(packages,library,character.only=T,quietly=T))
source("./lib/common_functions.R")
source("./lib/consistent_themes.R")
source("./scripts/GWAS/analysis/GWAS_analysis_scripts.R") 

# Set null device
set_null_device("png")

# Print
Sys.info()
sessionInfo()
```

# Load data

```{r} 
# Standard data
df <- readRDS("./data/dataset/df.RDS")
tr <- read.tree("./data/tree/tree.treefile")

## KPC Plasmids 
KPC_info_clusters <- readRDS("./data/KPC_plasmid/KPC_containing_clusters_mat.RDS") %>% as.data.frame 
KPC_containing_clusters <- KPC_info_clusters %>% `colnames<-`(paste0(colnames(.),"_KPC")) %>% {ifelse(.=="KPC Plasmid",1,0)} %>% as.data.frame%>% select_if(colSums(.)>0)
KPC_clusters <- colnames(KPC_containing_clusters) 
df <- left_join(df,KPC_containing_clusters %>% mutate(isolate_no = rownames(.))) 

# KPC_copy number
KPC_copy_number <- readRDS("./data/KPC_coverage/normalized/KPNIH1_KPC_plasmid_depth_by_feature_median_chromosome_normalized.RDS") %>% select(KPNIH1_RS28775_chr_median_norm) %>% rename(KPC_copy_number=KPNIH1_RS28775_chr_median_norm) %>% mutate(isolate_no = rownames(.))

df <- left_join(df,KPC_copy_number)
df$KPC_plasmid <- sapply(df$isolate_no,FUN=function(x){
  plasmid <- KPC_containing_clusters %>% subset(rownames(.)==x) %>% colSums(.) %>% subset(.>0)%>% names() %>% gsub("_KPC","",.)
  ifelse(length(plasmid)==0,"No Plasmid",plasmid)
}) %>% unlist

# KPC plasmid copy number
KPC_plasmid_copy_number <- readRDS("./data/KPC_plasmid_coverage/KPC_plasmid_copy_number_chromosome_normalized.RDS")
df <- left_join(df,KPC_plasmid_copy_number)

# KPC data 
KPC_data <- read.table("./data/KPC_blast/KPC_blast_master.txt",header = T)  
KPC_data$KPC_allele <- paste0("bla",str_split(KPC_data$sseqid,"\\|",simplify = T) %>% .[,ncol(.)])
df <- left_join(df,KPC_data %>% select(isolate_no,KPC_allele))
df$KPC_allele <- ifelse(is.na(df$KPC_allele),"No blaKPC",df$KPC_allele) 

df$KPC_copy_number_log_2 <- log2(df$KPC_copy_number)
df$KPC_plasmid_copy_number_log_2 <- log2(df$KPC_plasmid_median_copy_chromosome_normalized) 
```

## A. KPC copy number distribution

```{r}
A <- ggplot(df, aes(x= KPC_copy_number_log_2)) + geom_histogram(fill='black')  + theme_bw_me  + xlab("log<sub>2</sub>&nbsp;  <i>bla</i><sub>KPC</sub> copy number") + ylab("Isolates")  + geom_vline(xintercept = 1,linetype = 2,colour = 'brown')  + geom_vline(xintercept = 0,linetype = 2,colour = 'darkgray') + s_figure_9_format

A 
```

## B. blaKPC copy number

```{r}
df$KPC_allele <- factor(df$KPC_allele,levels = c(table(df$KPC_allele) %>% sort(decreasing = T) %>% names %>% subset(. != "No blaKPC") ,"No blaKPC"))

B <- ggplot(df, aes(x = KPC_allele, y = KPC_copy_number_log_2)) +
  geom_boxplot(outliers=F) +
  geom_jitter( height=0,
              size = 1, alpha = 0.5)  + theme_bw_me  + xlab("<i>bla</i><sub>KPC</sub> allele") + ylab("log<sub>2</sub>&nbsp; <i>bla</i><sub>KPC</sub> copy number") + geom_hline(yintercept = 1,linetype = 2,colour = 'brown')  + geom_hline(yintercept = 0,linetype = 2,colour = 'black')  + ylim(-3,5) + s_figure_9_format_plasmid_x

B
```

## C. KPC by copy number

```{r}
KPC_order <- df %>% group_by(KPC_plasmid) %>% summarise(med_KPC_CN_log2 = median(KPC_copy_number_log_2)) %>% arrange(-med_KPC_CN_log2) %>% subset(KPC_plasmid!='No Plasmid') %>% .$KPC_plasmid
df$KPC_plasmid <- factor(df$KPC_plasmid,levels = c(KPC_order,"No Plasmid"),labels = c(KPC_order,"No plasmid"))

C <- ggplot(df, aes(x = KPC_plasmid, y = KPC_copy_number_log_2)) +
  geom_boxplot(outliers=F) +
  geom_jitter(height=0,
              size = 1, alpha = 0.5)  + theme_bw_me  + xlab("<i>bla</i><sub>KPC</sub>-containing plasmid") + ylab("log<sub>2</sub>&nbsp; <i>bla</i><sub>KPC</sub> copy number")  + geom_hline(yintercept = 1,linetype = 2,colour = 'brown')  + geom_hline(yintercept = 0,linetype = 2,colour = 'black')     + ylim(-3,5) + s_figure_9_format_plasmid_x

C
```

## D. blaKPC plasmid copy number

```{r}
D <- ggplot(df, aes(x = KPC_plasmid, y = KPC_plasmid_copy_number_log_2)) +
  geom_boxplot(outliers=F) +
  geom_jitter( height=0, 
              size = 1, alpha = 0.5)  + theme_bw_me  + xlab("") + ylab("log<sub>2</sub> plasmid copy number") + geom_hline(yintercept = 1,linetype = 2,colour = 'brown')  + geom_hline(yintercept = 0,linetype = 2,colour = 'black')   + ylim(-3,5) + s_figure_9_format_plasmid_x

D
```

## E. Association between KPC and plasmid copy number

```{r}
E <-ggplot(df %>% subset(KPC_plasmid != "No Plasmid"), aes(x = KPC_plasmid_copy_number_log_2, y = KPC_copy_number_log_2)) + geom_point(size=1) + geom_abline(linetype = 2) +   theme_bw_me  + xlab("log<sub>2</sub> plasmid copy number") + ylab("log<sub>2</sub>&nbsp; <i>bla</i><sub>KPC</sub> copy number")  + geom_smooth(method = 'glm')  + ylim(-3,5)  + xlim(-3,5) + s_figure_9_format_plasmid_x

E
```

## F. KPC copy number by clade 

```{r}
f <- ggplot(df, aes(x = clade_I, y = KPC_copy_number_log_2)) +
  geom_boxplot(aes(fill=clade_I),outliers = F ) +
  geom_jitter(aes(color=clade_I), height = 0,
              size = 1, alpha = 0.5) + theme_bw_me  + xlab("") + ylab("log<sub>2</sub>&nbsp; <i>bla</i><sub>KPC</sub> copy number")  + clade_colors_scale_point + clade_colors_scale+ geom_hline(yintercept = 1,linetype = 2,colour = 'brown')  + geom_hline(yintercept = 0,linetype = 2,colour = 'darkgray') + ylim(-3,5) + s_figure_9_format
f
```

## G. blaKPC copy number

```{r}
G<- ggplot(df, aes(x = KPC_plasmid, y = KPC_copy_number_log_2, fill = clade_I)) +
  geom_boxplot(position = position_dodge(width = 0.8),outliers=F) +
  geom_jitter( 
              position = position_jitterdodge(jitter.width = 0.45, dodge.width = 0.8),
              size = 1, alpha = 0.5) +
  ggpubr::stat_compare_means() + theme_bw_me  + xlab("") + ylab("log<sub>2</sub>&nbsp; <i>bla</i><sub>KPC</sub> copy number") + geom_hline(yintercept = 1,linetype = 2,colour = 'brown')  + geom_hline(yintercept = 0,linetype = 2,colour = 'black') + clade_colors_scale   + ylim(-3,5)  + s_figure_9_format_plasmid_x

G
```

# Nearest neighbor analysis

```{r} 
isolates_w_plasmid <- subset(df,KPC_plasmid!="No Plasmid") %>% .$isolate_no
isolates_w_KPC <- subset(df,KPC_copy_number>0) %>% .$isolate_no
nn_dataset <-  df %>% select(isolate_no,KPC_plasmid,KPC_copy_number,blbli_dich_num) %>% subset(isolate_no %in% isolates_w_KPC)

phylogenetic_distance <- readRDS("./data/tree/phylo_dist.RDS") %>% as.data.frame %>% subset(rownames(.) %in% df$isolate_no) %>% select(any_of(df$isolate_no)) %>% subset(rownames(.) %in% isolates_w_KPC) %>% select(any_of(isolates_w_KPC)) %>% as.matrix
 
variant_distance <- readRDS("./data/nearest_neighbor_comparisons/pairwise_snp_dist_mat.RDS") %>% as.data.frame %>% select(any_of(isolates_w_KPC)) %>% subset(rownames(.) %in% isolates_w_KPC)  %>% as.matrix

plasmids <- colnames(KPC_containing_clusters)

nn_comps_all_vars_of_interest <- pbmcapply::pbmclapply(plasmids,function(variable_name){
  isolates_w_plasmid <- df[,c("isolate_no",variable_name)] %>% subset(get(paste(variable_name))==1) %>% select(isolate_no) %>% unlist
 variant_nn_data <- lapply(isolates_w_plasmid,FUN=nearest_neighbor_algorithm,tip_name_variable = 'isolate_no',phylogenetic_distance = phylogenetic_distance,variant_distance = variant_distance,annotate=T,metadata=df,variables_of_interest=c("clade_I","KPC_plasmid","KPC_copy_number",'blbli_dich_num',"MVB_num","IR_num"), comparison=T,comparison_feature = 'KPC_plasmid') %>% do.call(rbind,.)
  nn_comps_with_variant <- nearest_neighbor_analysis(variant_nn_data,binary_vars = c("blbli_dich_num"),categorical_vars = c("KPC_plasmid","clade_I"), continuous_vars = c("KPC_copy_number","MVB_num","IR_num"),log_2=T)%>% as.data.frame
  nn_comps_with_variant_sum <- nearest_neighbor_summary_statistics(nn_comps_with_variant,binary_vars = c("blbli_dich_num"),categorical_vars = c("KPC_plasmid","clade_I"), continuous_vars = c("KPC_copy_number","MVB_num","IR_num"),log_2=T)   
  lists <- list(variant_nn_data,nn_comps_with_variant,nn_comps_with_variant_sum) %>% `names<-`(c("variant_nn_data","nn_comps_with_variant","nn_comps_with_variant_sum")) 
  return(lists)
},mc.cores=7) %>% `names<-`(plasmids)

variant_stats <- lapply(nn_comps_all_vars_of_interest,FUN=function(x){x$nn_comps_with_variant}) %>% do.call(rbind,.)
summary_stats <-  lapply(nn_comps_all_vars_of_interest,FUN=function(x){x$nn_comps_with_variant_sum }) %>% do.call(rbind,.) %>% mutate(plasmid=names(nn_comps_all_vars_of_interest))
variant_stats$KPC_copy_number_change <- ifelse(variant_stats$KPC_copy_number_diff>0,"increase",ifelse(variant_stats$KPC_copy_number_diff<0,"decrease","neutral"))
``` 

## H. Change in copy number by plasmid type

```{r}
# Order the plasmids
variant_stats$isolate_comparison_value <- factor(variant_stats$isolate_comparison_value,levels = c(KPC_order,"No Plasmid"),labels = c(KPC_order,"No plasmid"))
# Scale
nearest_neighbor_scale <-  scale_color_manual(values = hues::iwanthue(length(unique(variant_stats$isolate_comparison_value))+1),name = "Nearest neighbor's <i>bla</i><sub>KPC</sub> plasmid",guide = guide_legend(nrow=3,title.position = "top", label.position = "right"))

H <- ggplot(data=variant_stats,aes(x=isolate_comparison_value,y=KPC_copy_number_log_2_diff)) + geom_boxplot(outliers=F) + 
  geom_jitter(height = 0,width = 0.25,alpha = 0.5,aes(color=nearest_neighbor_comparison_value),size=1) + 
  geom_hline(yintercept = 1,linetype = 2,colour = 'brown')  + geom_hline(yintercept = 0,linetype = 2,colour = 'black') + theme_bw_me +  nearest_neighbor_scale+ xlab("") + ylab("Change in log<sub>2</sub>&nbsp;<i>bla</i><sub>KPC</sub> copy number") + s_figure_9_format_plasmid_x
H
```

## I. Directional change in KPC copy number

```{r}
I <- ggplot(data=variant_stats,aes(x=isolate_comparison_value,fill=KPC_copy_number_change)) + geom_bar() + xlab("") + ylab("Isolates") + theme(legend.position = 'bottom') + theme_bw_me + scale_fill_manual(breaks = c("increase","decrease"),values = c("maroon","gray"),name = 'Change in <i>bla</i><sub>KPC</sub> gene copy number',labels = c("Increase","Decrease"), guide = guide_legend(nrow=3,title.position = "top", label.position = "right"))  +s_figure_9_format_plasmid_x
 
I_legend <- cowplot::get_legend(I) %>% cowplot::plot_grid()

I
```

## J. IR Change

```{r}
J <- ggplot(data=variant_stats,aes(x=isolate_comparison_value,y=IR_num_log_2_diff)) + geom_boxplot(outliers=F) + geom_jitter(height = 0,width = 0.25,alpha = 0.5,aes(color=nearest_neighbor_comparison_value),size=1)  + theme_bw_me + nearest_neighbor_scale + xlab("") + ylab("Change in log<sub>2</sub> IR MIC") + geom_hline(yintercept = 1,linetype = 2,colour = 'brown')  + geom_hline(yintercept = 0,linetype = 2,colour = 'black') + ylim(-6,6)  +s_figure_9_format_plasmid_x

J_legend <- cowplot::get_legend(J) %>% cowplot::plot_grid()

J
```

## K. MVB change

```{r}
K <- ggplot(data=variant_stats,aes(x=isolate_comparison_value,y=MVB_num_log_2_diff)) + geom_boxplot(outliers=F) + geom_jitter(height = 0,width = 0.25,alpha = 0.5,aes(color=nearest_neighbor_comparison_value),size=1)  + theme_bw_me +  nearest_neighbor_scale  + xlab("") + ylab("Change in log<sub>2</sub> MVB MIC") + geom_hline(yintercept = 1,linetype = 2,colour = 'brown')  + geom_hline(yintercept = 0,linetype = 2,colour = 'black') + ylim(-6,6)  + s_figure_9_format_plasmid_x

K
```

## L. BL/BLI change
```{r}  
L <- ggplot(data=variant_stats,aes(x=isolate_comparison_value,fill=blbli_dich_num_binary_diff)) + geom_bar() + xlab("") + ylab("Isolates")+ theme(legend.position = 'bottom') + theme_bw_me + scale_fill_manual(values = c("black","gray","green4"),breaks = c("gain","same",'loss'),labels = c("Acquisition","Same","Reversion"),name = "Change in BL/BLI resistance status",guide = guide_legend(ncol=1,title.position = "top", label.position = "right"))  + s_figure_9_format_plasmid_x

L_legend <- cowplot::get_legend(L) %>% cowplot::plot_grid()

L
```

# Curate full figure

```{r,fig.height=15,fig.width=12}
legends <- cowplot::plot_grid(I_legend,J_legend,L_legend,ncol=3,axis = 't')
figure <- cowplot::plot_grid(A + theme(legend.position = 'none'),B,C,D,E ,f + theme(legend.position = 'none'),G+ theme(legend.position = 'none'),H + theme(legend.position = 'none'),I+ theme(legend.position = 'none'),J+ theme(legend.position = 'none'),K+ theme(legend.position = 'none'),L+ theme(legend.position = 'none'),ncol=3,labels = "AUTO")
figure_w_legends <- cowplot::plot_grid(figure,legends,ncol=1,rel_heights = c(1,0.1))
figure_w_legends
```

## Save figure

```{r}
ggsave("./figures/s_figure_9_KPC_copy_number_dynamics.png",plot=figure_w_legends, units='in',height=15,width=12,dpi=900,bg='white')
```
