---
title: "curate_metadata_tree"
output: html_document
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
require(knitr)
knitr::opts_knit$set(echo = TRUE,warning = FALSE, message = FALSE)  
knitr::opts_knit$set(root.dir='~/Desktop/gl_mount/Analysis/Combination_resistance/pre-intro-blbli-resistance-ms/') 
```

# Environment
```{r}
# Packages
packages <- c("readxl","tidyverse","ape","phytools")
lapply(packages,library,character.only=T)

# Functions
source("./scripts/dataset_curation/metadata_curation_functions.R")

# Print environment
Sys.info()
sessionInfo()
``` 

# Clean the patient meadata
Source: UPENN metadata file (459_patients_data_for_posting.xls)
NFS Path: /nfs/turbo/umms-esnitkin/Project_Penn_KPC/Analysis/data/459_patients_data_for_posting.xls
## Load and remove isolate with incomplete data
```{r}
# Load in the patient data
metadata <- read_xls('../../data/459_patients_data_for_posting.xls') %>% as.data.frame
metadata$isolate_no <-  paste0("PCMP_H",metadata$isolate_no)
rownames(metadata) <- metadata$isolate_no 

# Remove patient with mismatched SENSITIRE isolate number
metadata <- metadata %>% subset(isolate_no != "PCMP_H473")
```

## Update spatiotemporal data
```{r}
metadata <- metadata %>% mutate(# Region
                                          Region = ifelse(Region == "Northern CA", "Southern CA",ifelse(
                                                          Region == "Houston","Texas", ifelse(
                                                          Region == "Tampa","Florida",ifelse(
                                                          Region == "Louisville","Kentucky","")))),
                                          # Recode LTACH Variable (Alphabetical Coding)
                                          LTACH_Recode = dplyr::recode(LTACH,
                                                                # Southern CA
                                                                EG =  "A", EH =  "B", EI =  "C", EJ =  "D", EL =  "E", GP =  "F",
                                                                GQ =  "G", GR =  "H", GS =  "I", LM =  "J", OG =  "K", SG =  "L",
                                                                # Houston
                                                                DO = "M", DP = "N", DQ = "O", IL = "P", IS = "Q", TL = "R",
                                                                # Tampa
                                                                FB = "S", FD = "T",
                                                                # Louisville
                                                                EZ = "U"),  
                                          LOSbeforeCx = dplyr::recode(LOSbeforeCx, '-202' = 202))

```

## Recode antibiotic exposure history data
```{r}
## Create new days of therapy (DOT) variables
metadata$AG_day <- metadata %>% select(amikacin_day,gent_day,tobra_day) %>% rowSums()
metadata$third_plus_ceph_day <- metadata %>% select(ceftazidime_day,ctx_day,ceftaroline_day,cefepime_day) %>% rowSums()
metadata$FQ_day <- metadata %>% select(cipro_day,levo_day) %>% rowSums() 
metadata$pseudoPCN_day <- metadata %>% select(zosyn_day,ceftazidime_day) %>% rowSums() 

## Binarize antibiotic exposure history (1 day was chosen)
DOT_vars <- colnames(metadata) %>% subset(grepl("day",.))
for(i in DOT_vars){
  metadata[,paste0(gsub("_day","",i),"_1d")] <-  binarize_abx_history(i,1,metadata)
}

## Binarized data
metadata$any_antibiotic_1d <- ifelse(rowSums(metadata %>% select(any_of(DOT_vars)))>0,1,0)
```

# Antibiotic resistance metadata
Source; Custom SENSITITRE broth microdilution assay
NFS Paths: /nfs/turbo/umms-esnitkin/Project_Penn_KPC/Analysis/Antibiotic_resistance/data/SENSITITRE MICs.xlsx
## Load data
```{r}
SENSITITRE <- read_excel("../../Antibiotic_resistance/data/SENSITITRE MICs.xlsx", range = "B1:N460") %>% as.data.frame
SENSITITRE <- SENSITITRE %>% mutate(isolate_no = paste0("PCMP_H",str_remove(gsub('CRKP_','',SENSITITRE$ISOLATES),"^0+"))) %>% `rownames<-`(.$isolate_no)

# PCMP_H472 & PCMP_H473 were excluded due to data reporting issues
SENSITITRE <- subset(SENSITITRE,!isolate_no %in% c("PCMP_H472","PCMP_H473"))
```

## Clean variables
```{r}
## Convert names
SENSITITRE <-  dplyr::rename(SENSITITRE,MVB = "MEV",IR = "IMK", CST = "COL", AMK = "AMI", FOS = "FOS+",PT = "P/T4")

## Fix MVB data due to miscoding (vaborbactom is a constant 8 µg/mL not 4 µg/mL)
table(SENSITITRE$MVB)
SENSITITRE <- SENSITITRE %>% mutate(MVB = dplyr::recode(MVB,"2/4" = "2/8"))

## Create MIC Variable
MIC_variables <- c("Ceftazidime-avibactam"="CZA","Meropenem-vaborbactam"="MVB","Imipenem-relebactam"="IR","Tigeycline"="TGC","Colistin"="CST","Amikacin"="AMK","Plazomicin"="PLZ","MERO","IMI")

## Numerize MIC Data
for(i in MIC_variables){
  SENSITITRE[,paste0(i,"_num")] <- SENSITITRE[,paste0(i)]  %>% gsub("≤","",.) %>% gsub(">","",.) %>% gsub("/.","",.) %>% trimws(.,'both') %>% as.numeric
}

## Factorize Raw MIC Data
for(i in MIC_variables){
  SENSITITRE[,i] <- factorize_MIC(SENSITITRE,i)
}

## Log-2 Transform Data
for(i in MIC_variables){
  SENSITITRE[,paste0(i,"_log_2")] <- log2(SENSITITRE[,paste0(i,"_num")])
}
```

## Convert Raw MIC data to phenotypic resistance categories
Select references:
1. https://academic.oup.com/cid/article/72/8/1484/5868454?login=true
2. https://www.sciencedirect.com/science/article/pii/S0255085721041062
```{r}
SENSITITRE <- SENSITITRE %>% mutate(
  ### Meropenem (MERO): ≤1/2/≥4
  MERO_cat = ifelse(MERO_num <= 1,"S",ifelse(MERO_num >= 4,"R","I")),
  ### Imipenem (IMI): ≤1/2/≥4
  IMI_cat = ifelse(IMI_num <= 1,"S",ifelse(IMI_num >= 4,"R","I")),
  ### Ceftazidime-avibactam: Susceptible (MIC≤8/4), Resistant (MIC≥16/4)  
  CZA_cat = ifelse(CZA_num <= 8,"S",ifelse(CZA_num >= 16,"R","")),
  ### Meropenem-vaborbactam (MVB): Susceptible (MIC≤4/8), Intermediate (MIC=8/8), Resistant (MIC≥16/8)  
  MVB_cat = ifelse(MVB_num <= 4,"S",ifelse(MVB_num == 8,"I","R")),
  ### Imipenem-relebactam (I-R): Susceptible (MIC≤1/4), Intermediate (MIC=2/4), Resistant (MIC≥4/4)  
  IR_cat = ifelse(IR_num <= 1,"S",ifelse(IR_num == 2,"I","R")),
  ### Tigecycline (TGC): Susceptible (MIC≤2), Intermediate (MIC=4), Resistant (MIC≥8)  
  TGC_cat = ifelse(TGC_num <= 2,"S",ifelse(TGC_num == 4,"I","R")),
  ### Colistin (CST): Intermediate (MIC≤2), Resistant (MIC≥4)
  CST_cat = ifelse(CST_num <= 2,"I",ifelse(CST_num >= 4,"R","")),
  ### Amikacin (AMK): Susceptible (MIC≤16), Intermediate (MIC=32), Resistant (MIC≥64)   
  AMK_cat =  ifelse(AMK_num <= 16,"S",ifelse(AMK_num == 32,"I","R")),
  ### Plazomicin (PLZ): Susceptible (MIC≤2), Intermediate (MIC=4), Resistant (MIC≥8) 
  PLZ_cat = ifelse(PLZ_num <= 2,"S",ifelse(PLZ_num == 4,"I","R")))

## Factorize Categorical Data
MIC_variables_cat <- paste0(MIC_variables,"_cat") 
for(i in MIC_variables_cat){
  SENSITITRE[,i] <- factor(as.vector(SENSITITRE[,i]),
                           levels = c("S","I","R"),
                           labels = c("Susceptible","Intermediate","Resistant"))
}

## Create Binary Non-Susceptible Variable
MIC_variables_no_CST <- c("MERO","IMI","MVB","IR","CZA","AMK","PLZ","TGC")
for(i in MIC_variables_no_CST){
  SENSITITRE[,paste0(i,"_dich")] <- binarize_MIC(SENSITITRE,paste0(i,"_cat"))
}
SENSITITRE$CST_dich <- ifelse(SENSITITRE$CST_cat == "Resistant","Non-Susceptible","Susceptible")

## Factorize Dichotomous Data
MIC_variables_dich <- paste0(MIC_variables,"_dich") 
for(i in MIC_variables_dich){
  SENSITITRE[,i] <- factor(as.vector(SENSITITRE[,i]),
                           levels = c("Susceptible","Non-Susceptible"),
                           labels = c("Susceptible","Non-Susceptible"))
}

## Numerical Dichotemous Variable for Resistance
for(i in MIC_variables_dich){
  SENSITITRE[,paste0(i,"_num")] <- ifelse(SENSITITRE[,i] == "Susceptible",0,1)
}
```

## Create BL/BLI combination therapy outcome
```{r}
SENSITITRE$blbli_dich <- ifelse(SENSITITRE$MVB_dich=="Non-Susceptible"|SENSITITRE$IR_dich=="Non-Susceptible","Non-Susceptible","Susceptible")  %>% factor(.,levels = c("Susceptible","Non-Susceptible"),labels = c("Susceptible","Non-Susceptible"))
SENSITITRE$blbli_dich_num <- ifelse(SENSITITRE$blbli_dich == "Non-Susceptible",1,0) 
```

# Kleborate data
Source: Kleborate data
NFS Path: /nfs/turbo/umms-esnitkin/Project_Penn_KPC/Analysis/Combination_resistance/pre-intro-blbli-resistance-ms/data/kleborate/klebsiella_pneumo_complex_output.txt
ENV: kleborate_v2_4_1.yml
## Load kleborate
```{r} 
kleborate <- read.delim(file='./data/kleborate/klebsiella_pneumo_complex_output.txt',sep='\t')
kleborate<- kleborate %>% mutate(isolate_no = gsub("_contigs_l1000","",strain)) %>% `rownames<-`(.$isolate_no) 
```

## Curate ST data
```{r}
# Clean ST Variable
kleborate <- kleborate %>% mutate(ST = gsub("ST","",ST) %>% gsub("-1LV","*",.),
                                  ST258 = ifelse(ST == "258",1,0))
```

# Create metadata dataframe
```{r}
# Merge metadata and sensititre data
df <- left_join(metadata,SENSITITRE,by="isolate_no") 
# Only keep ST258 isolates
df <- df %>% left_join(.,kleborate %>% select(isolate_no,ST,ST258))
df <- subset(df,ST258==1)
# Keep rows corresponding to the tree and match ordering (important for future work)
df <- df %>% `rownames<-`(df$isolate_no)    
```

# Phylogenetic tree data
Source: SNPKIT V2.0 run using CP008827 (KPNIH1) reference genome w/ PCMP_H243 as root 

## Load data
```{r}
tr <- read.tree(file="../../../Sequence_data/variant_calling/2025-02-17_snpkit_Penn_KPC_ST258/output_files/2025_02_23_21_11_49_core_results/phylokit/results/IQtree/Penn_KPC_ST258.treefile")  %>% 
                    drop.tip(
                    # Reference Sequence: https://www.ncbi.nlm.nih.gov/nuccore/CP008827
                    c('gi|661922017|gb|CP008827.1|','PCMP_H243','PCMP_H473'))

tr_midpoint = phytools::midpoint.root(tr) 
ape::write.tree(tr_midpoint,paste0("./data/tree/tree.treefile"))
```

## Generate phylogenetic distance data
```{r} 
phylo_dist_midpoint <- ape::cophenetic.phylo(tr_midpoint)
saveRDS(phylo_dist_midpoint,"./data/tree/phylo_dist.RDS") 
```

# Order dataframe
```{r} 
# Only keep ST258 isolates in tree
df <- subset(df,isolate_no %in% tr$tip.label) 
# Keep rows corresponding to the tree and match ordering (important for future work) 
df <- df %>%.[match(as.vector(tr$tip.label), .$isolate_no),]
```

# Get clade data
```{r}
library(ggtree)
root_number <-  tr_midpoint$edge[,1] %>% as.numeric %>% min
clade_nodes <-  tr_midpoint$edge[tr_midpoint$edge[,1] == root_number,] %>% .[,2]  
 
ggtree(tr_midpoint) + geom_point2(aes(subset=node==clade_nodes[1]), color='red', size=5)  + geom_point2(aes(subset=node==clade_nodes[2]), color='blue', size=5)
 
clade_1_descendants = tr_midpoint$tip.label[ phangorn::Descendants(x = tr_midpoint,node = clade_nodes[[1]],type='tips') %>% unlist]
clade_2_descendants = tr_midpoint$tip.label[ phangorn::Descendants(x = tr_midpoint,node = clade_nodes[[2]],type='tips') %>% unlist]

df$clade_I <- ifelse(df$isolate_no %in% clade_1_descendants,"Clade I","Clade II")
gheatmap(ggtree(tr_midpoint),df %>% select(clade_I))
```

# Write final dataframes
```{r} 
write.csv(df,file = "./data/dataset/df.csv")
saveRDS(df,file = "./data/dataset/df.RDS")
``` 