---
title: "results_section_X_MVB_rates"
author: "Kyle Gontjes"
date: "`r Sys.Date()`"
output: html_document
---


```{r setup, include=FALSE}
require(knitr)
opts_knit$set(echo = TRUE,warning = FALSE, message = FALSE,error=F,comment=NA) 
rm(list = ls())
opts_knit$set(root.dir='~/Desktop/gl_mount/Analysis/Combination_resistance/pre-intro-blbli-resistance-ms/')
```

# Environment

```{r environment, echo=F, message=FALSE, include=F, results=F}
#Packages
packages <- c("tidyverse","cowplot","ggtree","cowplot","ggtext",'lme4','sjPlot')

#Load Packages
invisible(lapply(packages,library,character.only=T,quietly=T))
source("./lib/consistent_themes.R")
source("./lib/common_functions.R")
```

# Load data

```{r}
df <- readRDS("./data/dataset/df.RDS")
df$clade_I_num <- ifelse(df$clade_I == "Clade I",1,0)

# Canonical
canonical <- readRDS("./data/carbapenem_resistance_panel/carbapenem_resistance_panel_matrix.RDS")
df <- left_join(df,canonical)

# tn4401 data
tn4401 <- readRDS("./data/TETyper/TETyper_curated.RDS")  
df <- left_join(df,tn4401)

# load the kleborate
kleborate <- readRDS("./data/kleborate/kleborate_curated.RDS") %>% as.data.frame
df <- left_join(df,kleborate) 

# KPC Plasmids 
KPC_info_plasmids <- readRDS("./data/KPC_plasmid/KPC_containing_plasmids_mat.RDS") %>% as.data.frame
KPC_info_clusters <- readRDS("./data/KPC_plasmid/KPC_containing_clusters_mat.RDS") %>% as.data.frame 
KPC_containing_clusters <- KPC_info_clusters %>% `colnames<-`(paste0(colnames(.),"_KPC")) %>% {ifelse(.=="KPC Plasmid",1,0)} %>% as.data.frame%>% select_if(colSums(.)>0)
KPC_clusters <- colnames(KPC_containing_clusters)
Non_KPC_clusters <- KPC_info_clusters %>% `colnames<-`(paste0(colnames(.),"_non_KPC")) %>% {ifelse(.=="Non-KPC plasmid",1,0)} %>% as.data.frame %>% select_if(colSums(.)>0)
df <- left_join(df,KPC_containing_clusters %>% mutate(isolate_no = rownames(.))) %>% left_join(Non_KPC_clusters %>% mutate(isolate_no = rownames(.)))  
df$KPC_plasmid <- lapply(df$isolate_no,FUN=function(x){
  KPC_containing_clusters[rownames(KPC_containing_clusters) == x,] %>% select_if(colSums(.)>0) %>% colnames() %>% {ifelse(length(.)==0,"None",paste0(.,collapse=" + "))}})  %>% unlist

# KC depth input
KPC_plasmid_depth <- readRDS("./data/KPC_coverage/normalized/KPNIH1_KPC_plasmid_depth_by_feature_median_chromosome_normalized.RDS")
KPC_plasmid_depth$isolate_no <- rownames(KPC_plasmid_depth)
                                         
KPC_gene_depth <- KPC_plasmid_depth %>% select(isolate_no,KPNIH1_RS28775_chr_median_norm)

df <- left_join(df,KPC_gene_depth)
```

## MVB experiment
```{r}
MVB_evolution_data <- read_csv("./data/evolution_experiment/curated_MVB_evolution_experiment_data.csv")
```

# Create annotated evolution dataset
```{r}
MVB_evolution_data <- MVB_evolution_data %>% left_join(.,df %>% select(isolate_no,MVB,KPNIH1_RS28775_chr_median_norm,clade_I,clade_I_num,PBP_any,OmpK36_L3_mutations,OmpK36GD,OmpK36TD,OmpK36_truncation_kleborate,OmpK36_non_syn,OmpK36_promoter_IS,tn4401_allele,KPC_present,KPC_type,KPC_plasmid,AA018_KPC,AA552_non_KPC))
MVB_evolution_data$category <- MVB_evolution_data %>% select(clade_I,KPC_type,OmpK36_L3_mutations,AA018_KPC,AA552_non_KPC) %>% apply(.,1,FUN=function(x){paste0(c(x["clade_I"],x["KPC_type"],names(x %>% subset(. %in% c(1)))),collapse=" + ")}) %>% gsub("OmpK36_L3_mutations",'Loop 3 insertion',.) %>% gsub("AA552_non_KPC","AA552 non-KPC plasmid",.) %>% gsub("AA018_KPC","AA018 KPC plasmid",.)  %>% gsub("+ No KPC +","",.) %>% gsub("\\+\\+","\\+",.)
MVB_evolution_data$MVB_ratio_log10 <- log10(MVB_evolution_data$MVB_ratio)
MVB_evolution_data$AA018_KPC_AA552_nonKPC <- ifelse(MVB_evolution_data$AA018_KPC==1 & MVB_evolution_data$AA552_non_KPC==1,1,0)
MVB_evolution_data$cens <- ifelse(MVB_evolution_data$over_LoD, "none", "left")
MVB_evolution_data$cens <- factor(MVB_evolution_data$cens, levels = c("none", "left"))
```


# Across
```{r}
MVB_evolution_data_avg <- MVB_evolution_data %>% subset(over_LoD==T) %>% group_by(isolate_no) %>% mutate(ratio_avg = mean(MVB_ratio))  %>% distinct
```

# Question A - Differences across isolates  
```{r}
paste0("Overall modeling")
lm(MVB_ratio_log10~isolate_no ,data=MVB_evolution_data) %>% anova
paste0("Overall modeling w/ over LoD specific")
lm(MVB_ratio_log10~isolate_no,data=MVB_evolution_data%>% subset(over_LoD==T)) %>% anova
``` 

# Question B - Differences between clades?  
```{r}
paste0("Overall modeling")
lmer(MVB_ratio_log10~ 1 + clade_I_num + (1 | isolate_no),data=MVB_evolution_data) %>% tab_model()
paste0("Overall modeling LOD")
lmer(MVB_ratio_log10~ clade_I_num + (1 | isolate_no),data=MVB_evolution_data%>% subset(over_LoD==T) ) %>% tab_model()
```

# Question C: Impact of loop 3?
```{r}
paste0("Overall modeling")
lmer(MVB_ratio_log10~OmpK36GD + (1 | isolate_no),data=MVB_evolution_data) %>% tab_model()
paste0("Overall modeling in clade I")
lmer(MVB_ratio_log10~OmpK36GD + (1 | isolate_no),data=MVB_evolution_data%>% subset(clade_I=="Clade I")) %>% tab_model()
paste0("Overall modeling > LoD")
lmer(MVB_ratio_log10~OmpK36GD + (1 | isolate_no),data=MVB_evolution_data%>% subset(over_LoD==T) ) %>% tab_model()
paste0("Overall modeling >LoD in clade I")
lmer(MVB_ratio_log10~OmpK36GD + (1 | isolate_no),data=MVB_evolution_data%>% subset(over_LoD==T & clade_I=="Clade I") ) %>% tab_model()
```

# Question D: Impact of AA018 plasmid 
```{r}
paste0("Overall modeling")
lmer(MVB_ratio_log10~AA018_KPC + (1 | isolate_no),data=MVB_evolution_data) %>% tab_model()
paste0("Overall modeling in clade I")
lmer(MVB_ratio_log10~AA018_KPC + (1 | isolate_no),data=MVB_evolution_data%>% subset(clade_I=="Clade I") ) %>% tab_model()
paste0("Overall modeling > LOD")
lmer(MVB_ratio_log10~AA018_KPC + (1 | isolate_no),data=MVB_evolution_data%>% subset(over_LoD==T) ) %>% tab_model()
paste0("Overall modeling > LOD in clade I")
lmer(MVB_ratio_log10~AA018_KPC + (1 | isolate_no),data=MVB_evolution_data%>% subset(over_LoD==T & clade_I=="Clade I") ) %>% tab_model()
```

# Question E: Presence of both AA018 + AA552 plasmid
```{r}
paste0("Overall modeling")
lmer(MVB_ratio_log10~AA018_KPC_AA552_nonKPC + (1 | isolate_no),data=MVB_evolution_data) %>% tab_model()
paste0("Overall modeling in clade I")
lmer(MVB_ratio_log10~AA018_KPC_AA552_nonKPC + (1 | isolate_no),data=MVB_evolution_data%>% subset(clade_I=="Clade I") ) %>% tab_model()
paste0("Overall modeling LOD")
lmer(MVB_ratio_log10~AA018_KPC_AA552_nonKPC + (1 | isolate_no),data=MVB_evolution_data%>% subset(over_LoD==T) ) %>% tab_model()
paste0("Overall modeling LOD in clade I")
lmer(MVB_ratio_log10~AA018_KPC_AA552_nonKPC + (1 | isolate_no),data=MVB_evolution_data%>% subset(over_LoD==T & clade_I=="Clade I") ) %>% tab_model()
```

# Question F: Plasmid and loop 3 modeled  
```{r}
paste0("Overall modeling")
lmer(MVB_ratio_log10~OmpK36GD+AA018_KPC_AA552_nonKPC + (1 | isolate_no),data=MVB_evolution_data) %>% tab_model()
paste0("Overall modeling in clade I")
lmer(MVB_ratio_log10~OmpK36GD+AA018_KPC_AA552_nonKPC + (1 | isolate_no),data=MVB_evolution_data%>% subset(clade_I=="Clade I") ) %>% tab_model()
paste0("Overall modeling LOD")
lmer(MVB_ratio_log10~OmpK36GD+AA018_KPC_AA552_nonKPC + (1 | isolate_no),data=MVB_evolution_data%>% subset(over_LoD==T) ) %>% tab_model()
paste0("Overall modeling LOD in clade I")
lmer(MVB_ratio_log10~OmpK36GD+AA018_KPC_AA552_nonKPC + (1 | isolate_no),data=MVB_evolution_data%>% subset(over_LoD==T & clade_I=="Clade I") ) %>% tab_model()

```

# Question G: Plasmid interaction with loop 3
```{r}
paste0("Overall modeling")
lmer(MVB_ratio_log10~OmpK36GD*AA018_KPC_AA552_nonKPC + (1 | isolate_no),data=MVB_evolution_data) %>% tab_model()
paste0("Overall modeling in clade I")
lmer(MVB_ratio_log10~OmpK36GD*AA018_KPC_AA552_nonKPC + (1 | isolate_no),data=MVB_evolution_data%>% subset(clade_I=="Clade I") ) %>% tab_model()
paste0("Modeling > LOD")
lmer(MVB_ratio_log10~OmpK36GD*AA018_KPC_AA552_nonKPC + (1 | isolate_no),data=MVB_evolution_data%>% subset(over_LoD==T) ) %>% tab_model()
paste0("Modeling > LOD in clade I")
lmer(MVB_ratio_log10~OmpK36GD*AA018_KPC_AA552_nonKPC + (1 | isolate_no),data=MVB_evolution_data%>% subset(over_LoD==T & clade_I=="Clade I") ) %>% tab_model()
```

# Question H: KPC copy number
```{r}
paste0("Overall modeling")
lmer(MVB_ratio_log10~KPNIH1_RS28775_chr_median_norm + (1 | isolate_no),data=MVB_evolution_data) %>% tab_model()
paste0("Overall modeling in clade I")
lmer(MVB_ratio_log10~KPNIH1_RS28775_chr_median_norm + (1 | isolate_no),data=MVB_evolution_data%>% subset(clade_I=="Clade I") ) %>% tab_model() 
paste0("Modeling > LOD in clade I")
lmer(MVB_ratio_log10~KPNIH1_RS28775_chr_median_norm + (1 | isolate_no),data=MVB_evolution_data%>% subset(over_LoD==T)) %>% tab_model()
paste0("Modeling > LOD in clade I")
lmer(MVB_ratio_log10~KPNIH1_RS28775_chr_median_norm + (1 | isolate_no),data=MVB_evolution_data%>% subset(over_LoD==T & clade_I=="Clade I") ) %>% tab_model()
```

# Question I: L3 + KPC
```{r}
paste0("Overall modeling")
lmer(MVB_ratio_log10~OmpK36GD+KPNIH1_RS28775_chr_median_norm + (1 | isolate_no),data=MVB_evolution_data) %>% tab_model()
paste0("Overall modeling in clade I")
lmer(MVB_ratio_log10~OmpK36GD+KPNIH1_RS28775_chr_median_norm + (1 | isolate_no),data=MVB_evolution_data%>% subset(clade_I=="Clade I") ) %>% tab_model() 
paste0("Modeling > LOD")
lmer(MVB_ratio_log10~ OmpK36GD+ KPNIH1_RS28775_chr_median_norm + (1 | isolate_no),data=MVB_evolution_data%>% subset(over_LoD==T)) %>% tab_model()
paste0("Modeling > LOD in clade I")
lmer(MVB_ratio_log10~OmpK36GD + KPNIH1_RS28775_chr_median_norm + (1 | isolate_no),data=MVB_evolution_data%>% subset(over_LoD==T & clade_I=="Clade I"))%>% tab_model()
```

# Question J: Loop 3 insertion with KPC copy number  
```{r}
paste0("Overall modeling")
lmer(MVB_ratio_log10~OmpK36GD*KPNIH1_RS28775_chr_median_norm + (1 | isolate_no),data=MVB_evolution_data) %>% tab_model()
paste0("Overall modeling in clade I")
lmer(MVB_ratio_log10~OmpK36GD*KPNIH1_RS28775_chr_median_norm + (1 | isolate_no),data=MVB_evolution_data%>% subset(clade_I=="Clade I") ) %>% tab_model() 
paste0("Modeling > LOD")
lmer(MVB_ratio_log10~ OmpK36GD*KPNIH1_RS28775_chr_median_norm + (1 | isolate_no),data=MVB_evolution_data%>% subset(over_LoD==T)) %>% tab_model()
paste0("Modeling > LOD in clade I")
lmer(MVB_ratio_log10~OmpK36GD*KPNIH1_RS28775_chr_median_norm + (1 | isolate_no),data=MVB_evolution_data%>% subset(over_LoD==T &clade_I=="Clade I"))  %>% tab_model()
```

# Full model
```{r}

paste0("Overall modeling")
lmer(MVB_ratio_log10~OmpK36GD+AA018_KPC_AA552_nonKPC+ KPNIH1_RS28775_chr_median_norm + (1 | isolate_no),data=MVB_evolution_data) %>% tab_model()
paste0("Overall modeling in clade I")
lmer(MVB_ratio_log10~OmpK36GD+AA018_KPC_AA552_nonKPC+ KPNIH1_RS28775_chr_median_norm + (1 | isolate_no),data=MVB_evolution_data%>% subset(clade_I=="Clade I") ) %>% tab_model() 
paste0("Modeling > LOD")
lmer(MVB_ratio_log10~ OmpK36GD+AA018_KPC_AA552_nonKPC+KPNIH1_RS28775_chr_median_norm + (1 | isolate_no),data=MVB_evolution_data%>% subset(over_LoD==T)) %>% tab_model()
paste0("Modeling > LOD in clade I")
lmer(MVB_ratio_log10~OmpK36GD+AA018_KPC_AA552_nonKPC+KPNIH1_RS28775_chr_median_norm + (1 | isolate_no),data=MVB_evolution_data%>% subset(over_LoD==T &clade_I=="Clade I"))  %>% tab_model()
```


# TLDR:

1. Positive association with clade

2. Positive association with Loop 3

3. Positive association with having AA018 KPC

4. No association having both AA018 w/ KPC and AA552 without KPC

5. Interaction modeling didn't have important findings